% Created 2024-11-20 Wed 07:37
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{preamble}
\author{Adithya Nair}
\date{\today}
\title{23MAT204}
\hypersetup{
 pdfauthor={Adithya Nair},
 pdftitle={23MAT204},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.4 (Org mode 9.7.11)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\part{Mathematics}
\label{sec:orgdc81465}
\chapter{Revision Linear Algebra}
\label{sec:orgbe3b617}
Every matrix satisfies its own characteristic matrix.

What this means is

\begin{align}
\lambda^3 - 2\lambda^2 + \lambda -4 &= 0 \\
A^3 - 2A^2 + A -4I &= 0 \\
A^2 - 2A + I - 4A^{-1} &= 0 \\
A^{-1} &= \frac{1}{4}[A^2 - 2A + I] \\
X = A^{-1}b &= \frac{1}{4}[A^2 - 2A + I]b
\end{align}
There are some implications behind this. Here's a problem, generate a random 3x3 matrix, find the ranks of
\((A- \lambda_1 I)\) and then \((A-\lambda_1 I)(A-\lambda_2 I\)

What you will see is that the rank reduces upon multiplying roots of the
characteristic equation
\section{Ways To Calculate Whether A Matrix Is Positive Definite}
\label{sec:org7f1112d}
\begin{enumerate}
\item The minors are positive then negative alternating.
\item The eigenvalues are positive.
\end{enumerate}
\section{Large Eigenvalue Computation}
\label{sec:org1922ad5}
Large eigenvalues are computed by using a matrix multiplication method.
This method involves:
\section{Specral Decomposition}
\label{sec:org573c85c}
\[S = Q \Lambda Q^T = \lambda_1 q_1 q_1^T + \lambda_2 q_2 q_2^T + \cdots + \lambda_n q_n q_n^T\]

This allows us to represent a matrix by the sum of n rank 1 matrices.
This is used in square symmetric matrices.

This is the basis behind Principal Component Analysis.
\section{Singular Value Decomposition}
\label{sec:orgba59127}
For rectangular matrices.
$$A = U \Sigma V^T  = \sigma_1 u_1 v_1^T + \sigma_2u_2v_2^T + \cdots + \sigma_r u_r v_r^T$$

This is mainly used for getting useful properties about the matrix
without needing to perform significant operations on this matrix, such
as the orthogonality.

Singular values are the eigenvalues of a matrix which are \textbf{non-zero}

So a rectangular matrix \(m \times n\) would have \(AA^T\) which is
\(m \times m\) and \(A^TA\) which is \(n \times n\). The number of
singular values is whether \(m\) or \(n\) is lower.

The trace of \(A^TA\) equal to the sum of all \(a_{ij}^2\) The trace is
the sum of all eigenvalues of \(A^T A\), and For \(A_{m \times n}\)
that's the sum of the eigenvalues squared.

This is known as the \textbf{Frobenius Norm}

This method can also be used to generate orthonormal bases for the 4
Fundamental Subspaces
\section{The Geometry of SVD}
\label{sec:orgd624ebd}
Since we have an orthogonal matrix, diagonal and orthogonal matrix\ldots{}
It's a rotation step followed by a stretching step, finally ended by
another rotation step.
\section{Polar Decomposition}
\label{sec:org32559cc}
This decomposition is a special case of the Singular Value
Decomposition. Where you decompose the elements into the polar form with
an orthogonal matrix and a positive semi-definite matrix.

For 2D, \[\vec{x} = r(\cos(\theta) + sin(\theta))\]

\begin{aligned}
    A &= U \Sigma V^T \\ 
      &= U(V^TV) \Sigma V^T \\
      &= (UV^T)(V\Sigma V^T)\\
      &= Q \times S \\
\end{aligned}

Here S is the scaling matrix, and Q is the orthogonal matrix.

Although my assumptions might be wrong, this might be a way to generate
rotation matrices for higher-dimension spaces. Might be useful in
exploring the semantic spaces of LLMs.
\section{Principal Component Analysis}
\label{sec:org9e53a99}
Why do we use normalization while working with data? We use
normalization to ensure that the units are eliminated while working with
that given data. \textbf{This normalization is done by subtracting by the
average and divide by the standard deviation.}

Matrix where the row and columns sums are equal.
\section{Norms Of Vectors And Matrices}
\label{norms-of-vectors-and-matrices}
Norms are a way to measure the size of a vector/matrix. The norm of a
vector which calculates the magnitude is known as the \(L^2\) norm or
the Euclidean norm \(\|v\|_2\). This number gives us the length of the
vector.

In general, the \(\|u\|_p\) or the p-norm of a vector is
\([|u_1|^p + |u_2|^p + \cdots + |u_n|^p]^{\frac{1}{p}}\).

Now, the \(L_1\) norm would be the sum of the components of the vector.
This is the sum of the projections to each corresponding axis.

The \(L_3\) norm would be
\([|u_1|^3 + |u_2|^3 + \cdots + |u_n|^3]^{\frac{1}{3}}\)

The \(L_\infty\) norm would be
\([|u_1|^\infty + |u_2|^\infty + \cdots + |u_n|^\infty]^{\frac{1}{\infty}}\).
This returns the maximum vector component of the vector.

Let's take the equation, \(\|v\|_1 = 1\) \begin{aligned}
    x + y \&= 1 \\
    x - y \&= 1 \\
\begin{itemize}
\item x + y \&= 1 \\
\end{itemize}
    -x - y = 1 \\
\end{aligned}

\label{fig:l1eq1}
\url{./figures/l1eq1}

The S-norm of a vector \(\vec{x}\) is \(\vec{x}^T S \vec{x}\). When S is
a symmetric positive definite matrix, this S-norm is known as the energy
of vector \(\vec{v}\)

There are three types of Matrix norms:

\begin{enumerate}
\item Spectral Norm

\item Frobenius Norm =
\(\sqrt{\sigma_1^2 + \sigma_2^2 + \cdots + \sigma_r^2}\)

\item Nuclear Norm
\end{enumerate}
\begin{enumerate}
\item Spectral Norm
\label{sec:org9c34ba0}
We know that the vector norm for a vector \(\vec{x}\) is nothing but
\(\vec{x}^T \vec{x}\). We take this property.

\begin{aligned}
    Max \|A\|_2^2 &= Max \frac{\|Ax\|_2^2}{\|x\|_2^2} \\
              &= Max \frac{x^TA^TAx}{x^Tx} \\
              &= Max \{\lambda_i(S)\} = \lambda_1 = \sigma_1^2
\end{aligned}
\item Frobenius Norm
\label{frobenius-norm}
The Frobenius norm for a matrix M, \(\begin{bmatrix}
    a_{11} & a_{12}\\
a_{21} & a_{22} \\
\end{bmatrix}\) is the equation
\(\sqrt{a_{11}^2 + a_{12}^2 + a_{21}^2 + a_{22}^2}\)
\item Nuclear Norm
\label{nuclear-norm}
The nuclear norm is the sum of the singular values of a matrix A.

For an identity matrix,

\begin{itemize}
\item The Spectral Norm is 1

\item The Frobenius Norm is \(\sqrt{n}\)

\item The Nuclear Norm is \({n}\)
\end{itemize}

For an orthogonal matrix,

\begin{itemize}
\item The Spectral Norm is 1

\item The Frobenius Norm is \(\sqrt{n}\)

\item The Nuclear Norm is \({n}\)
\end{itemize}
\end{enumerate}
\section{Best Low Rank Matrix}
\label{best-low-rank-matrix}
We say that a matrix is the best approximation of another matrix, based
on the Frobenius Norm.. For a singular value decomposition
\(A = U\Sigma V^T\), if we assume that the singular values are arranged
in descending order\ldots{} We can select the singular value range where the
values are significant contributors to the final matrix.

We can then reduce the size of \(U, \Sigma\) and \(V^T\) into a smaller
matrix B, based on the number of singular values chosen which would give
the best approximation.

Let \(A = U\Sigma V^T\) where
\(\Sigma: \sigma_1 \geq \sigma_2 \geq \cdots \sigma_n\), then B =
\(U_{m\times m} \Sigma V^T{n \times n}\) is a best rank-k approx. to A.
Where, S is a diagonal matrix of \(n \times n\) where
\(s_i = \sigma_i (i = 1\cdots k)\) else \(s_i = 0\), by best B is a
solution to \(min_B \|A - B\|F\) where rank(B) = k
\chapter{Multi variable Optimization}
\label{sec:orgf8250c5}
We minimize \(f(x_1,x_2,\cdots,x_3)\).
\section{Contour Curves}
\label{sec:orgaf4f197}
\section{Multi variable Calculus}
\label{sec:orgd8d8ba6}
The points where \(\nabla f = \vec{0}\) are called the stationary points. The Hessian matrix should be positive definite for a minima and negative definite for a maxima. A point where there is no change, is known as a saddle point.

\begin{align*}
f(x,y) = x^3 + y^3 + 2x^2 + 4y^2 + 6 \\
\frac{\partial f}{\partial x} = 3x^2 + 4x \\
\frac{\partial^2 f}{\partial x} = 6x + 4 \\
\frac{\partial f}{\partial y} = 3y^2 + 8y\\
\frac{\partial f}{\partial y\partial x} = 0\\
\frac{\partial^2 f}{\partial y} = 6y + 8\\
x = 0,\frac{-4}{3} \\
y = 0,\frac{-8}{3} \\
\end{align*}
\section{Newton's Method}
\label{sec:org1169c54}
Newton's Method is a numerical method to find a minimum of a function.

Iterative formula:

\[
x_{n+1} = x_n - \frac{f'(x_n)}{f''(x_n)}
\]

Find the minimum of \(f(x) = x^2 + \frac{54}{x}\) using Newton's method.

Newton's method can be practically done for all functions by evaluating the first and second derivatives numerically and then apply the formula

$$x_{n+1} = x_n - \frac{f(x_n)}{f'(x_n)} \text{to find root of $f(x)$ = 0}$$
$$x_{n+1} = x_n - \frac{f(x_n)}{f'(x_n)} \text{to find minimum of $f(x)$(root of $f'(x) = 0)}$$

Evaluation of first and second derivatives numerically:

$$f'(a) = \frac{f(a+\Delta a)-f(a-\Delta a)}{2 \Delta a}$$
$$f''(a) = \frac{f(a+\Delta a)- 2f(a) + f(a-\Delta a)}{(\Delta a)^2}$$
\section{How solution to a linear system can be found as a solution of an optimization problem.}
\label{sec:orgb2dd519}
Find an objective function whose solution is specified as \(\vec{x} : A\vec{x} = b\)

Take for example, \(\frac{1}{2}x^T A x - b^T x + c\), \(x \in R^n, A = A^T\)
\section{Numerical Algorithm Of Gauss-Jacobi Method}
\label{sec:orgb1272e0}

Input A = [a\textsubscript{ij}], X0 = x\textsuperscript{(0)}, tolerance TOL, maximum number of iterations,

\(Ax = b\)

We separate A,
A = D - L - U

$$(D-L-U)x = b$$
$$(Dx= b + (L+U)x$$
$$(x= D^{-1}b + D^{-1}(L+U)x$$

$$x^{(k)} = Tx^{x^{(k-1)}} + c$$


The code to implement the method:
\begin{verbatim}
A = [5,-2,3;-3,9,1;2,-1,-7];
b = [-1;2;3];

n = 100; % No. of iterations
D = diag(diag(A));
L = -tril(A,-1);
U = -triu(A,1);

X = zeros(size(A)(1),n);
T = inv(D)*(L+U);
c = inv(D)*b;

for i = 2:n
  X(:,i) = T*X(:,i-1) + c;
  if(X(:,i) == X(:,i-1))
    i-1
    break;
  end
end
\end{verbatim}
\section{Gauss-Siedel Iteration Method}
\label{sec:org75af681}
The iterative method is somewhat similar to <Gauss-Jacobi method>.

The only difference is that the new values are computed by already existing new values.

\begin{verbatim}
A = [5,-2,3;-3,9,1;2,-1,-7];
b = [-1;2;3];

n = 100; % No. of iterations
D = diag(diag(A));
L = -tril(A,-1);
U = -triu(A,1);

X = zeros(size(A)(1),n);
T = inv(D-L)*U;
c = inv(D-L)*b;

for i = 2:n
  X(:,i) = T*X(:,i-1) + c;
  if(X(:,i) == X(:,i-1))
    display("The iteration converges at:")
    i-1
    break;
  end
end
\end{verbatim}

$$(D-L)x^{k} = Ux^{k-1} + b$$
$$x^{k} = (D-L)^{-1} \vec{b} + (D-L)^{-1}(U \vec{x})$$

If the matrix is 'diagonally dominant', where the magnitude of the diagonal elements should be greater than the sum of the magnitude of the other elements(absolute value) in the same row.
\section{Unidirectional/Line search}
\label{sec:org0633b41}
For a given function,

Find the minima of \(f(x) = (x-1)^2 + (y-2)^2\) starting from \((0,0)\) along the direction of x-axis.

How to perform uni-directional search:
\begin{itemize}
\item Write the parametric representation of the line search, $$\vec{s}(t) = \vec{a} + t\vec{b}$$
\item Write the function in terms of \(t\), \(f(\vec{s}(t))\). this is a single variable function.
\item Find the minimum of \(f(\vec{s}(t))\), \(t^*\)
\item Obtain the solution for unidirectional search by substituting \(t^*\) in \(\vec{s}(t)\)

Find the minimum of the function \(f(x) = (x-2)^2 - y\), starting from the point \((-1,0)\) along \((1,1)^T\)
\end{itemize}
\section{Directions Of Change}
\label{sec:orge08a33e}
A direction given by the vector \$x\textsuperscript{(k)}\} is a descent direction only if the function value decreases along that direction from the point, \(x^{(k)}\), when
$$\nabla f(x^{(k)}\cdot d < 0$$

A direction given by the vector \$x\textsuperscript{(k)}\} is an ascent direction only if the function value increases along that direction from the point, \(x^{(k)}\), when
$$\nabla f(x^{(k)}\cdot d > 0$$

Question: Check whether the given function have a descent direction \(x = (2,-1)\) along the given directions \(d_1\) and \(d_2\)
$$f = 2x_1^{2} + x_2^2-2x_1x_2 + 2x_1^3 + x_2^4$$
\section{Directions Of Descent/Ascent In Numerical Algorithms.}
\label{sec:orgdca814d}
In any numerical algorithm to find the optimum of an unconstrained problem, the iterative formula used is:
$$x^{k+1} = x^k + \alpha^kd^k$$
Where \(\alpha^k\) is the step length in step k and \(d^k\) is the direction of descent in step k, if it's a minimization problem or a direction of ascent if it's a maximization problem.

Newton's Method: \(x_{k+1} = x_k -(H(x_k)^{-1}\nabla f(x_k))\)
Gradient DEscent - \(x_{k+1} = x_k + \alpha_kd_k\) where \(d_k = - \nabla f(x_k)\) and \(\alpha =\) using line search

A direction along which the function increases rapidly from a point is given by the gradient of the function at that point
\section{Steepest Descent}
\label{sec:orga0ba941}

A direction along which the function increases rapidly from a point is given by the gradient of the function at that point

A direction along which the function decreases rapidly from a point is given by the negative of the gradient of the function at that point

The iterative formula of method of steepest descent is:

$$x^{k+1} = x^k + \alpha^k d^k$$

The descent direction \(d^k\) is along the steepest descent direction, \(d^k = - \nabla f(x^k\). The step length \(\alpha^k\) is obtained by performing a unidirectional search from \(x^k\) along the direction \(d^k\), by minimizing,

This method produces successive directions that are perpendicular to each other.

Near the minima, during line search the convergence is very slow.
\section{Numerical Method}
\label{sec:org9078ef9}
To implement this programmatically, we can write\ldots{}
For \(Ax=b\), we can convert this to an equivalent optimization problem.
$$f(x) = \frac{1}{2}x^TAx + - b^Tx + c$$

We have a linear system of equations, which we can be solved by \(Ax=b\)

We can write \(g_0 = Ax_{0} - b\)
\section{Conjugate Gradient Descent}
\label{sec:orgca2a159}
The problem with normal gradient descent is that there is no way to predict when the function converges. There's a method that can reliably find the convergence in a predictable fashion. That's the conjugate gradient method.

First we must understnand the conjugate direction and the terms associated with them
\begin{enumerate}
\item Krylov Subspace
\label{sec:orgf7db9a0}
Let \(A \in R^{n \times n}, b \in R^n\), the Krylov subspace \(K_j(A,b)\) is defined as \(K_j(A,b) = span \{\vec{b}, \vec{Ab}, A^2b, A^3b, \dots , A^{j-1}b\}\). Thus \(K_j(A,b)\) is subspace of \(\mathbb{R}^n\)
\item Krylov Matrix
\label{sec:org7231627}
The Krylov Matrix is just the matrix consisting of the basis vectors as the column vectors. The Krylov subspace is the column space of such a matrix.
\item Motivation For Krylov Subspaces
\label{sec:orgd1f6a68}
We know the Cayley Hamilton theorem, and its use in calculating the inverse from the characteristic equation
Interestingly, the calculation for the inverse leads to a linear combination of the basis vectors of the Krylov subspace when solving linear systems.
\item Conjugate Directions
\label{sec:org1c6e865}
For A real symmetric matrix \(n \times n\) with rank \(n\)

The directions \(d_0 d_1 \dots d_{n-1}\) are said to be A-conjugate if,

$$d_i Ad_j = 0 $$ for \(i \neq j\)

It's a new definition for orthogonality(not orthonormal)
\item Numerical Method
\label{sec:orgdc2c3e1}
Input \(A,b,x^{0}\)

Compute \(\vec{r_{0}} = b - A x^0\)
\begin{verbatim}
A = [1 2 3;2 3 4;3 4 5];

b = [6 9 12]';
x = randi([-9,9], length(b),1);
r = b - A*x;
d = r;
alpha = (r'*r)/(d'*(A*d));
beta = 0;

for i = 1:size(A,1)
  x = x + alpha*d;
  rnew = r - alpha*(A*d);
  beta = (rnew'*rnew)/(r'*r);
  d = rnew + beta*d;
  r = rnew;
  alpha = (r'*r)/(d'*(A*d));
  if sqrt(rnew) < 1e-10
    break;
    end
end

x
r
\end{verbatim}
\end{enumerate}
\chapter{Probability}
\label{sec:org357cbfd}
\section{Probability mass function}
\label{sec:orge00aee5}
For discrete inputs
The probabilty mass function for a discrete random variable \(x\) is defined as, \(f(x) = P(X = x)\)
\begin{enumerate}
\item Properties
\label{sec:org7358a59}
\begin{itemize}
\item \(0 \leq f(x) \leq 1\)
\item \(\Sigma f(x) = 1\)
\item \(p(\text{at least n outcomes}) = p(x \geq n) = f(n) + \cdots + f(end)\)
\end{itemize}
\item Mean Of Random Variable
\label{sec:org4d31728}
$$U_x = E(x) = \Sigma x f(x)$$
\item Variance Of Random Variable
\label{sec:org41b4479}
$$\sigma_x^{2} = E((x-\mu_{x}^2)) = \Sigma (x-\mu)^2 f(x) = \Sigma x^2f(x) - \mu^{2}$$
$$\sigma_x^{2} = E(X^2) - \mu^2$$
\end{enumerate}
\section{Cumulative Distribution Function}
\label{sec:orgf73ed55}
A cumulative distribution function is just adding the previous terms to the new terms.
\section{Probability density function - For continuous inputs}
\label{sec:orgc0ba6f0}
For continuous inputs

The probability density function for a continuous random variable \(x\) is defined as, \(P(a \le x \le b) = \int_a^b f(x)dx\)
\begin{enumerate}
\item Binomial Distribution
\label{sec:orge753019}
The Bernoulli Process

\begin{enumerate}
\item Repeated trails,
\item Each trial results in an outcome that may be classified as a success
\item The trials are independent.
\end{enumerate}

Where n is the number of trials, x is the random variable distribution, \(q = 1 - p\)
\(^nC_xp^{x} q^{n-x}\)

We have,
\begin{align*}
^nC_x \text{Different ways $x$ successes can occur among $n$ trials} \\
p^x - \text{probability of getting $x$ successes} \\
(1-p)^{n-x} - \text{probability of getting $n-x$ failures}
\end{align*}
\item Cumulative Distribution Function
\label{sec:orgac0a093}
$$\int_0^x f(y)dy$$
\item Mean Of Random Variable
\label{sec:orgc8bae5e}
$$U_x = E(x) = \Sigma x f(x)$$
\item Properties
\label{sec:org584fc46}
\begin{itemize}
\item \(f(x) \ge 0\)
\item \(\int_xf(x) = 1\), where \(\int_x\) is the range for which \(x\) is defined.
\item For a continuous random variable, the probability at a point is \(0\), the probability that \(x\) takes a discrete value is zero.
\end{itemize}
\item Uniform Distribution
\label{sec:org66c1cf4}
A probability distribution which is constant throught the interval, given by the function,

$$f(x) = \frac{1}{b-a}  $$

$$Mean = \frac{a+b}{2}$$
$$Variance = \frac{(b-a)^2}{12}$$

$$F(x) = \int_{-\infty}^xf(x)dx = \int_{-\infty}^x \frac{1}{b-a}  $$

$$
f(X) =    \begin{cases}
0, - \infty < x < 0 \\
x, 0<x<1 \\
1, x > 1 \\
     \end{cases}
$$
\item Variance Of Random Variable
\label{sec:orga75e792}
$$\sigma_x^{2} = E((x-\mu_{x}^2)) = \Sigma (x-\mu)^2 f(x) = \Sigma x^2f(x) - \mu^{2}$$

$$\sigma_x^{2} = E(X^2) - \mu^2$$
\end{enumerate}
\section{Joint Probability Distribution}
\label{sec:org53f1470}
This is given for two variables, a function \(f(x,y)\) gives us a probability, \(P(X= x_1, X = x_2\)
\begin{enumerate}
\item Properties
\label{sec:org740cf9a}
\begin{itemize}
\item \(0 \le f(x_1,y_1) \le 1\)
\item \(\Sigma_x \Sigma_y f(x,y) = 1\)
\end{itemize}
\item Marginal Distribution Function
\label{sec:org1aec55e}
Marginal distribution functions are the probability distributions of the variables themselves.

\begin{itemize}
\item The distribution of \(x_1\) is given by the columns, and \(x_2\) is given by the rows. In other words you're fixing the values and adding the remaining.
\end{itemize}
\item Mean Of Marginal Values
\label{sec:org903badc}

\(Mean(X_1) = E(X_1) = \Sigma_{x1} x_1 f_{1)(x_{1})\)
\end{enumerate}
\section{Double Integrals}
\label{sec:orga53d345}
\(\int_x \int_y f(x,y) dy dx\)
\begin{enumerate}
\item Area Of A Region
\label{sec:orgdbdf7eb}
$$R= \int_R 1 dx dy$$
\end{enumerate}
\section{Joint Probability Density Function}
\label{sec:org5424b9f}
For a continuous scalar field, \(P(a \le x  \le b, c \le y \le d)\)
\begin{enumerate}
\item Properties
\label{sec:orgcb39de0}
\begin{itemize}
\item \(\int\int f(x,y) dx dy = 1\)
\end{itemize}
\end{enumerate}
\section{Independent Probabilities}
\label{sec:org44b59e9}

Two events \(A\) and \(B\) are said to be independent if \(P(A|B) = P(A)\) or \(P(B|A) = P(B)\)

Two discrete random variables \(x_1\) and \(x_2\) are said to be independent if \(f(x_1,x_2)\) = \(f(x_1)f_2(x_2) \forall x_1 \& x_2\)
\section{Binomial Distribution}
\label{sec:org394c5ce}
The Bernoulli process has these properties:
\begin{enumerate}
\item The experiment consists of repeated trials
\item Each trial results in an outcome that may be classified as a success or failure.
\end{enumerate}
\section{Poisson Distribution}
\label{sec:org12e322a}
$$\lim_{n \rightarrow \infty} P(X=x) = \frac{e^{-\lambda}\lambda^x}{x!}$$
\section{Continuous Uniform Distribution}
\label{sec:org0c516dc}
$$f(x) = \frac{1}{b-a}$$
\section{Exponential Distribution}
\label{sec:orgd733d12}
Let N be a random variable with a Poisson Distribution with mean \(\lambda x\) and X be the random variable that gives the time between each arrival or count. Then,


$$P(X > x) = P(N=0) = \frac{e^{-\lambda x}(\lambda x)^0}{0!} = e^{-\lambda x}$$
PDF,
$$\lambda e^{-\lambda x}$$
\section{Gaussian/Normal Distribution}
\label{sec:orgd01d597}
\begin{itemize}
\item It is one of the most common probability distribution

$$f(x) = \frac{1}{\sqrt{2 \pi} \sigma} e^{\frac{-(x-\mu)^2}{2 \sigma^2}}$$
\end{itemize}
\section{Covariance And Correlation Coefficient}
\label{sec:org1323c4c}

Correlation is a measure of how much two data points depend on each other.

$$\rho = \frac{Cov(X_1,X_2)}{\sigma_1 \sigma_2}$$
\section{Multiple Joint Distributions}
\label{sec:org2542038}

\begin{itemize}
\item Multiple discrete random variables

Multinomial distributions, $$P(X_1 = x_{1}, X_2 = x_2, ... , X_k = x_k) = \frac{n!}{x_1! x_2! x_3! ... x_k!} p_1^{x_1}p_2^{x_2}$$
\end{itemize}
\chapter{Constrained Optimization}
\label{sec:org1a78328}
\section{Graphical Method}
\label{sec:org872ba81}
For a linear system, the optimum solution will always be at the boundary of the feasible region. Such systems will always have \textbf{either} -
\begin{enumerate}
\item Unique solutions
\item Infinitely many solutions
\item Unbounded solutions(solutions for a feasible region which is not fully closed.)
\end{enumerate}
\section{Analytical Solution}
\label{sec:org6b85623}
There are three problems which can be solved analytically
\begin{itemize}
\item Least squares problems
\item Linear optimization problems
\begin{itemize}
\item Convex optimization problems
\end{itemize}
\end{itemize}
\section{Convex Sets}
\label{sec:orgab68688}
Let S be a set and let \(x_1\) and \(x_2\) be elements of the set. If the line segment joining \(x_1\) and \(x_2\) is also an element of S. then we say that S is convex.

Mathematically:
\begin{itemize}
\item Let \(x_1\), \(x_2 \in S\)
\item If \(\alpha x_1 + (1 - \alpha) x_{2}\), then S is a convex set.
\end{itemize}
\begin{enumerate}
\item Examples
\label{sec:org8218b5c}
\begin{itemize}
\item The empty set
\item The singleton set.
\item Any vector space
\item Any line segment is a convex set.
\item Solution set of any linear equations
\item A halfspace is convex(a halfspace is defined by a linear inequality)
\item A polyhedron and polytope is convex
\end{itemize}
\end{enumerate}
\section{Convex Functions}
\label{sec:org58ee57b}
A function \(f: R^n \rightarrow R\) is a convex function if:
\begin{enumerate}
\item the domain of \(f\) is convex and
\item \(f(\theta x + (1 - \theta)y) \le \theta f(x) + (1- \theta) f(y)\)
\end{enumerate}
\section{Concave Function}
\label{sec:orgbc5c6e3}
A function \(f: R^n \rightarrow R\) is a convex function if:
\begin{enumerate}
\item the domain of \(f\) is convex and
\item \(f(\theta x + (1 - \theta)y) \le \theta f(x) + (1- \theta) f(y)\)
\end{enumerate}
\section{Second derivative test for convexity/concavity}
\label{sec:orge2b8734}

A twice differentiable function \(f(x)\) is convex if its Hessian \(\nabla^{2}f(x)\) is positive definite.
A twice differentiable function \(f(x)\) is concave if its Hessian \(\nabla^{2}f(x)\) is negative definite.
\section{Operations preserving convexity}
\label{sec:orge2ae2c6}
\begin{itemize}
\item A scalar multiple
\item Sum of convex Functions
\end{itemize}
\section{Relation between convex sets and convex functions}
\label{sec:org98eb5e4}

If \(g(x)\) is a convex function, then \(g(x) \le 0\) is a convex set.

An optimization problem in the standard form:
  Minimize \(f(x)\) subject to \(g_i(x) \le 0, i = 1,2,...,m\), \(h_i(x) = 0, i = 1,2,...,p\)

is said to be convex if

\begin{enumerate}
\item the objective function is convex
\item the inequality constraint functions, \(g_i(x)\) are convex
\item the equality constraint functions, h\textsubscript{j}(x) are linear.
\end{enumerate}
\section{Langrangian Multiplier}
\label{sec:org78a6849}
For an optimization problem,
At an optimal point \(x\), $$\nabla f(x^{*}) = \lambda \nabla g(x^{**})$$

We define a Langrangian function,
$$L = f(x) + \lambda g(x)$$
\section{Kuhn-Tucker Conditions}
\label{sec:orgefa5e45}
To recap, a standard optimization problem is, Minimize \(f(x_1,x_2,\cdots x_n}\) subject to \(g_1(x_1,\cdots , x_n) \le 0\)

These conditions are known as the Kuhn Tucker Conditions, the sufficient condition is that the problem is a convex optimization problem.

We write the Langrangian for a problem,

$$L = f + \lambda_{1} g_{1} + \lambda_2g_2 + \cdots \lambda_m g_m + \mu_1 h_1 \mu_2h_2 + \cdots \mu_p h_{p}$$


The first condition is that,

$$\frac{\partial L}{\partial x_i} = 0$$

This gives us \(n\) equations.

The second condition is that,

$$\lambda_i g_i(x) = 0$$

This gives us \(m\) equations

The third condition is,

$$\lambda_i \ge 0$$

This ensures that the sign does not flip, since they're paired with inequality constraints.

This gives us \(m\) equations

The fourth condition

$$g_i \le 0, h_j = 0$$

No. of variables = \(n+m+p\)

No. of KT conditions = \(n + 3m + p\)

No. of equality conditions = \(n + m + p\)
\part{Partial Differential Equations}
\label{sec:org1cc2e93}

When a function is dependent on only one variable is expressed using ODEs, when there are multiple variables, we have partial differential equations.

These include spatial co-ordinates with respect to time.

\textbf{Boundary Conditions} - Value of the dependent variable with respect to the independent variable at a certain point.

A differential equation is said to be non-linear if the equation has terms which are non-linear with respect to the dependent variable.


A differential equation is said to be linear if the dependent variable and its derivatives are linear in every term of the differential equation i.e. if there are no products or non-linear functions of the dependent variable and its derivatives.
\chapter{Linear Second Order PDEs}
\label{sec:org4e0f084}
For two independent variables, the general form of the pde is,

$$A \frac{\partial^2u}{\partial x^2}+B \frac{\partial ^2 u }{\partial x \partial y} + C \frac{\partial^2u}{\partial y^2}+ D = 0$$

For \(B^2 - 4AC\),:
\begin{itemize}
\item < 0, Elliptic equations
\item = 0, Parabolic equations
\item > 0, Hyperbolic equations
\end{itemize}
\chapter{Finite Difference Formula}
\label{sec:org81b33ee}

For a function T depending on x and y $$T(x,y)$$: The finite difference approximations of the partial derivatives,

$$\frac{\partial T}{\partial x} = \frac{T(x+h,y)-T(x,y)}{h}$$
$$\frac{\partial T}{\partial y} = \frac{T(x,y+h)-T(x,y)}{h}$$
\chapter{Laplace's Equation}
\label{sec:org5050931}

$$u_{xx} + u_{yy} = 0$$

We had an iterative formula for single-variable calculus,

\begin{align*}
y' &= f(x) \\
y(x+h) &= y(x) + hf(x) \\
\end{align*}

\begin{align*}
y' &= f(x) \\
y(x+h) &= y(x) + hf(x) \\
\end{align*}

The equation for evaluating these functions is,

$$T(x,y)  = \frac{T(x+h,y) + T(x-h,y) + T(x,y+h) + T(x,y-h)}{4}$$

This way, by writing the equation for every grid point, we obtain a system of linear equations, that can be solved by the Gauss-Siedel method, to obtain the value of the dependent variable at the grid points.
\part{Syllabus For Quiz 3}
\label{sec:orgd6158cb}
\begin{enumerate}
\item Convex optimization problem
\item Lagrangian Multiplier
\item Kuhn-Tucker Conditions
\item PDEs
\end{enumerate}
\end{document}
